## Capture-recapture data {#dataCMR}

```{r globalGetDataCMR, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r librariesDataCMR, echo = FALSE}
library(getWBData)
library(tidyverse)
library(lubridate)
library(validate)
library(rlang)
library(stringi)
```

```{r}
getNew_encounterHistories <- TRUE
```


### Get data
West Brook electrofishing data formatted for Capture-Mark-Recapture analysis for tagged individuals
```{r CMR}
source('./rForSourcing/envFunctions.R')

if(getNew_cdWB_CMR0) {
  cdWB_CMR0 <- 
    createCoreData(
      sampleType = "electrofishing", #"stationaryAntenna","portableAntenna"),
      whichDrainage = "west",
      columnsToAdd =
        c("sampleNumber",
          "river",
          "riverMeter",
          "survey",
          "pass",
          'observedLength',
          'observedWeight')
    ) %>%
    addTagProperties(
      columnsToAdd = 
        c("cohort",
          "species",
          "dateEmigrated",
          "sex",
          "species")
      ) %>%
    dplyr::filter(!is.na(tag), 
                  area %in% c("trib","inside","below","above"), 
                  !is.na(sampleNumber) 
                  ) %>%
    createCmrData(maxAgeInSamples = 20, 
                  inside = F, 
                  censorDead = F, 
                  censorEmigrated = T) %>% # may want to change censorEmigrated = T to = F
    addSampleProperties() %>%
    addEnvironmentalDaily() %>%
    addEnvironmentalInterval() %>%
    addKnownZ() %>%
    fillSizeLocation(size = F) #assumes fish stay in same location until observed elsewhere

  save(cdWB_CMR0, file = './data/cdWB_CMR0.RData')
  
} else {
  load(file = './data/cdWB_CMR0.RData')
}
str(cdWB_CMR0)
```

### Make encounter history files
```{r enc hist}
# https://stackoverflow.com/questions/69583424/using-tidy-eval-for-multiple-arbitrary-filter-conditions
# Assumes LHS is the name of a variable and OP is
# the name of a function
op_call <- function(op, lhs, rhs) {
  call(op, sym(lhs), rhs)
}

ehFilter <- function(data, cols, ops, vals) {
  exprs <- purrr::pmap(list(ops, cols, vals), op_call)
  data %>% dplyr::filter(!!!exprs)
}

# var is the variable to put in the encounter history (e.g. 'enc' or 'temp')
# occasionVar is the variable used to define samlping occasions (e.g. 'sampleNumber' or 'ageInsamples')
getEHDataWide <- function(d, cols, ops, vals, var, occasionVar = "ageInSamples", maxOccasionValue, valuesFill = 0){   
  d %>%
    ehFilter(cols, ops, vals) %>% 
    filter(ageInSamples > 0, get(occasionVar) <= maxOccasionValue) %>%
    #arrange(eval(substitute(occasionVar)), tag) %>%
    arrange(get(occasionVar)) %>% #need this to get correct order of columns. eval(substitute(occasionVar)) doesn;'t work 
    pivot_wider(
      id_cols = tag,
      names_from = eval(substitute(occasionVar)),
      names_prefix = substr(eval(substitute(var)), 1, 1), #"s",
      values_from = eval(substitute(var)),
      values_fill = valuesFill
    )
}

# occasionVar is the variable used to define sampling occasions (e.g. sampleNumber or ageInsamples)
getEH <- function(d, cols, ops, vals, occasionVar = "ageInSamples", maxOccasionValue){
  encWide <- getEHDataWide(d, cols, ops, vals, "enc", occasionVar, maxOccasionValue)
  eh <- as.matrix(encWide %>% select(-tag), nrow = nrow(encWide), ncol = ncol(encWide) - 1)
  
  flowWide <- getEHDataWide(d, cols, ops, vals, "meanFlow", occasionVar, maxOccasionValue)
  flowMatrix <- as.matrix(flowWide %>% select(-tag), nrow = nrow(flowWide), ncol = ncol(flowWide) - 1)
  
  temperatureWide <- getEHDataWide(d, cols, ops, vals, "meanTemperature", occasionVar, maxOccasionValue)
  temperatureMatrix <- as.matrix(temperatureWide %>% select(-tag), nrow = nrow(temperatureWide), ncol = ncol(temperatureWide) - 1)

  tags <- encWide %>% select(tag)
  
  data <- d %>%
    ehFilter(cols, ops, vals) %>% 
    filter(ageInSamples > 0, get(occasionVar) <= maxOccasionValue) %>%
    arrange(eval(substitute(occasionVar)), tag)
  
  cohorts <- tags %>% left_join(data %>% select(tag, cohort) %>% unique()) %>% select(cohort) 
  
  return(list(eh = eh, flow = flowMatrix, temperature = temperatureMatrix, tags = tags, cohorts = cohorts, data = data ))
}


if (getNew_encounterHistories) {

  # read down through the cols, ops, vals variables for filter conditions
  cols <- list( "river")
  ops <- list("==")
  vals <- list("wb obear")
  
  eh <- getEH(cdWB_CMR0, cols, ops, vals, occasionVar = "sampleNumber", maxOccasionValue = 200)
  fileName = paste0("eh_", stri_paste_list(vals, collapse = "_"))
  save(eh, file = paste0('./models/dataOut/', fileName, '.RData'))
  
  # read down through the cols, ops, vals variables for filter conditions
  cols <- list("cohort", "river", "sampleNumber")
  ops <- list("==", "==", "<")
  vals <- list(2003, "wb obear", 40)
  
  eh <- getEH(cdWB_CMR0, cols, ops, vals, occasionVar = "sampleNumber", maxOccasionValue = 200)
  fileName = paste0("eh_", stri_paste_list(vals, collapse = "_"))
  save(eh, file = paste0('./models/dataOut/', fileName, '.RData'))

    # read down through the cols, ops, vals variables for filter conditions
  cols <- list("cohort", "river", "sampleNumber")
  ops <- list("%in%", "==", "<")
  vals <- list(2003:2005, "wb obear", 400)
  
  eh <- getEH(cdWB_CMR0, cols, ops, vals, occasionVar = "sampleNumber", maxOccasionValue = 200)
  fileName = paste0("eh_", stri_paste_list(vals, collapse = "_"))
  save(eh, file = paste0('./models/dataOut/', fileName, '.RData'))

  
      # read down through the cols, ops, vals variables for filter conditions
  cols <- list("cohort", "river")
  ops <- list("%in%", "==")
  vals <- list(2002:2006, "wb obear")
  
  eh <- getEH(cdWB_CMR0, cols, ops, vals, occasionVar = "ageInSamples", maxOccasionValue = 10)
  fileName = paste0("eh_", stri_paste_list(vals, collapse = "_"))
  save(eh, file = paste0('./models/dataOut/', fileName, '.RData'))

}


```


### Wrangle data
```{r}

```

### Explore data
```{r}

ggplot(cdWB_CMR0 %>% filter(enc ==1), aes(year)) +
  geom_bar() +
  facet_grid(river + season ~ species)
  


```


### CMR metadata
#### adapted from https://github.com/Conte-Ecology/westBrookData/blob/master/getWBData/vignettes/westBrookDataIntro.Rmd

#### Column explanations
__tag__ PIT tag number, unique identifier, character

__cohort__ year the fish was born, assigned based on size at initial capture and size distributions of fish of known age

__detectionDate__ mostly self explanatory, but filled in for unobserved fish as the median capture date for all observed fish.

__sampleName__ An ordered identifier for sampling mostly for recognition by people who did the sampling. This is not very clean because early in the study samples were not taken at strictly seasonal intervals. sampleNumber is probably more useful and intuitive.

__sampleNumber__ A tidier identifier for samples that strictly increases by one for each season (4/yr)

__river__ River the fish was observed in. NA if the fish was not observed.

  __west brook__ The mainstem

  __wb jimmy__ Larger tributary that fish can move back and forth into from WB section 31 (Open Large from Letcher et al 2015)

  __wb mitchell__ Smaller tributary that fish can move back and forth into from WB section 35 (Open Small from Letcher et al 2015)

  __wb obear__ Smaller tributary that has a waterfall at its mouth, so fish can only move downstream into WB section 20 (Isolated Small from Letcher et al 2015)

__section__ Identifier for the 20m section that the fish was captured in. This is ordered from downstream to upstream starting at 1 within each river. 

__area__ inside = section 1:47 in the west brook, trib = tributary (not west brook), below = sections below inside sections, above = sections above the inside sections

__observedLength__ in mm

__survey__ shock = electroshocking survey

__pass__ electrofishing pass. 1 or 2 in the west brook (inside), 1 in tribs

__observedWeight__ in g wet weight

__species__  
bkt = brook trout (native, self-sustained population)  
bnt = brown trout (non-native, self-sustained population)  
ats = atlantic salmon (stocked through 2005, no reproduction)  

__dateEmigrated__ date of emigration from inside/tribs if observed to emigrate. Coded as emigrated if last observation was on PIT antenna or captured below or above 

__sex__ NA = unknown, f = female, m = male, p = precocious male (salmon only)

__enc__  
Logical, was the fish observed? (1 = yes, 0 = no)

__ageInSamples__ number of seasons since summer of the year of birth (1st summer = 0)

__sampleIndex__ sampleNumber rescaled to start at 1 and end at length(unique(sampleNumber)) for ease of looping in JAGS

__tagIndex__ ordered, unique individual identifier 1:N

__year__ of sample

__season__ 1 = spring, 2 = summer, 3 = fall, 4 = winter

__proportionSampled__ Occasionally the sample was not complete (e.g., skipped west brook but did the tributaries). This is the proportion of sections in the river of capture that were sampled.

__lagDetectionDate__ detection date lagged back one observation

__meanTemperature__ mean temperature between observation dates. If individual was not observed, median observation date for the sampling occasion was used.

__meanFlow__ mean flow between observation dates. If individual was not observed, median observation date for the sampling occasion was used.

__knownZ__ z is alive state, so this is '1' between first and last capture, and NA otherwise, unless the fish was known to be dead (e.g. tagging mortality or observed dead) in which case the value is set to '2'. There is also an option in the addKnownZ() function to useAntenna. This is useAntenna = FALSE by default, but could be set to TRUE to set knownZ to 1 up to the last antenna observation.



