## Electrofishing data {#dataElectro}

This section retrieves electrofishing data, including both tagged and untagged fish. Fish were untagged if they were too small (< 60 mm, 2 g wet weight) or were captured outside of the core study area (tributaries and 47 sections of the mainstem West Brook).
```{r globalGetDataElectro, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r librariesDataElectro, echo = FALSE}
# library(devtools)
# install_github(repo = "Conte-Ecology/westBrookData", subdir = "getWBData")
library(getWBData)
library(tidyverse)
library(lubridate)
#library(validate)

# devtools::install_github('bletcher/getPrepareWBData')
# use this if command above doesn't work: options(download.file.method = "wininet")

library(getPrepareWBData) # this has functions for preparing West Brook data
```

### Get data
```{r cdWB_electro0}

# default values for createCoreData()
# function (sampleType = "electrofishing", baseColumns = T, 
#    columnsToAdd = NULL, includeUntagged = F, whichDrainage = "west") 

if(getNew_cdWB_electro0) {
  cdWB_electro0 <- createCoreData(
        sampleType = "electrofishing",  #"stationaryAntenna","portableAntenna"
        columnsToAdd = c("sampleNumber",
                         "river",
                         "survey",
                         "pass",
                         "observedLength",
                         "observedWeight",
                         "comments"),
        includeUntagged = TRUE,
        whichDrainage = "west"
      ) %>%
      addTagProperties(
        columnsToAdd = c("cohort",
                         "species",
                         "dateEmigrated",
                         "sex",
                         "species"
        )
      ) %>%
      dplyr::filter(species %in% c( "bkt","bnt","ats"),
                    area %in% c("trib","inside","below","above"),
                    !is.na(sampleNumber)) %>%
      addSampleProperties() %>%
      addEnvironmental()
  
  save(cdWB_electro0, file = './data/cdWB_electro0.RData')
  
} else {
  load(file = './data/cdWB_electro0.RData')
}
str(cdWB_electro0)
```

### Wrangle data


```{r globalWrangle, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

This section takes cdWB_electro0 and cleans it to create cdWB_electro.  

```{r}
reclean_cdWB_electro <- FALSE
```


```{r wrangleElectroMain, cache = TRUE}

if(reclean_cdWB_electro){
  drainage <- 'west'

  # functions in getPrepareWBData library
  cdWB_electro <- cdWB_electro0 %>%
    cleanData(drainage) %>%
    mergeSites(drainage) %>%
    addNPasses(drainage) %>%
    mutate(drainage = drainage)
  
  save(cdWB_electro, file = './data/cdWB_electro.RData')
  
} else {
  load(file = './data/cdWB_electro.RData')
}
```

### Explore data

```{r lw, cache = TRUE}

ggplot(cdWB_electro, aes(observedLength, observedWeight, color = species)) +
  geom_point(alpha = 0.1) +
  scale_x_log10() +
  scale_y_log10() +
  theme_publication() +
  facet_wrap(~ species)

lwReg <- cdWB_electro %>%
  nest_by(species) %>%
  mutate(reg = list(lm(log(observedWeight) ~ log(observedLength), data = data)))

lwReg %>% summarise(broom::tidy(reg))  
lwReg %>% summarise(broom::glance(reg))
```



