## Flow effects on survival (phi) models {#modelCMR_Flow_4rivers}

The goal of this modelling exercise is to evaluate the effect of new tributary-specific stream flow estimates on survival of brook trout and brown trout. We will compare survival across the WB and tributaries with flow input data as 1) single flow estimate for all locations (historical approach) and 2) hindcasted flows for each tributary based on new tributary-specific flows which are available since 2000.

The goal is to find the best structure for the survival model, then compare survival estimates with tributary-specific flow to estimates with common flow across locations.

Structure options include
[species, cohort, season, isYOY, flow, flow^2]


```{r globalModelsNimble, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r librariesModelsNimble, echo = FALSE}
library(getWBData)
library(lubridate)
library(kableExtra)
library(GGally)
library(nimble)
library(nimbleEcology)
library(MCMCvis)
library(tidyverse)
```

```{r}
rerunSurivalModels <- FALSE
plotMCMCOutput <- TRUE
```


```{r}
#all the data
#load('./models/cmrFlow4rivers/dataOut/eh_2002200320042005200620072008200920102011201220132014.RData')
#load('./models/cmrFlow4rivers/dataOut/eh_200620072008.RData')

#250 fish from each cohort
#load('./models/cmrFlow4rivers/dataOut/eh_2002200320042005200620072008200920102011201220132014_n250.RData')
#250 fish from each of 3 cohorts
load('./models/cmrFlow4rivers/dataOut/eh_200620072008_n250.RData')

table(eh$data$cohort, eh$data$river)
```


### Model phiT_pT_psiT  
Phi and p and psi vary by time    
Includes transition probabilities among rivers
Added cohort (c) indexing to psi's
Fixed c,t indexing, change to t,c
Added ilogis() to betaPhi
Added hier vars on psi

#### Set up and run model
```{r phiT_pT_psiT_River2}

# Following https://oliviergimenez.github.io/bayesian-cr-workshop/worksheets/5_demo.html
if (rerunSurivalModels) {
  
  y <- eh$eh * eh$riverN
  (nCohorts <- nrow(unique(eh$cohorts)))
  (nSeasons <- nrow(unique(eh$seasons)))
  (nRivers <- length(unique(eh$data$riverN)))# rivers 1:4
  seasonArray <- c(3,4,1,2,3,4,1,2,3,4,1,2)
             

  first <- eh$first #apply(y, 1, function(x) min(which(x !=0)))
  last <- eh$last
  cohort = ((eh$cohorts) - min(eh$cohorts) + 1)$cohort #can't be a data frame or tibble
  
  zinits <- y + 1 # non-detection -> alive
  zinits[zinits == 2] <- 1 # dead -> alive
  zInitsNA <- ifelse(is.na(eh$flow), NA, 1)
  
  #  For 3 states, we are using 4
  # -------------------------------------------------
  # Parameters for 3 sites (A, B, C):
  # phiA: survival probability site A
  # phiB: survival probability site B
  # phiC: survival probability site B
  # psiAA: movement probability from site A to site A (reference)
  # psiAB = psiA[1]: movement probability from site A to site B
  # psiAC = psiA[2]: movement probability from site A to site C 
  # psiBA = psiB[1]: movement probability from site B to site A
  # psiBB: movement probability from site B to site B (reference)
  # psiBC = psiB[2]: movement probability from site B to site C
  # psiCA = psiC[1]: movement probability from site C to site A
  # psiCB = psiC[2]: movement probability from site C to site B
  # psiCC: movement probability from site C to site C (reference)
  # pA: recapture probability site A
  # pB: recapture probability site B
  # pC: recapture probability site C
  # -------------------------------------------------
  # States (z):
  # 1 alive at A
  # 2 alive at B
  # 3 alive at C
  # 4 alive at D
  # 5 dead
  # Observations (y):  
  # 1 not seen
  # 2 seen at A 
  # 3 seen at B
  # 4 seen at C
  # 5 seen at D
  # river names: c("west brook" = 1, "wb jimmy" = 2, "wb mitchell" = 3, "wb obear" = 4)
  # -------------------------------------------------
  
  ## model code using DHMMo distribution
  hmm.phiT_pT_psiT_DHMM <- nimbleCode({
      delta[1] <- 0.4                  # Pr(alive t = 1 and in river 1) = 0.4
      delta[2] <- 0.2
      delta[3] <- 0.2
      delta[4] <- 0.2
      delta[5] <- 0                    # Pr(dead t = 1) = 0
  
      for (r in 1:nRivers){
          betaPhiRiver[r] ~ dnorm(0,1)
          betaPRiver[r] ~ dnorm(0,1)
          
          betaPsiRiverA[r] ~ dnorm(0,1)
          betaPsiRiverB[r] ~ dnorm(0,1)
          betaPsiRiverC[r] ~ dnorm(0,1)
          betaPsiRiverD[r] ~ dnorm(0,1)
          
          betaPhiRiverOut[r] <- 1/(1 + exp(-betaPhiRiver[r]))
          betaPRiverOut[r] <- 1/(1 + exp(-betaPRiver[r]))
          
          betaPsiRiverOutA[r] <- 1/(1 + exp(-betaPsiRiverA[r]))
          betaPsiRiverOutB[r] <- 1/(1 + exp(-betaPsiRiverB[r]))
          betaPsiRiverOutC[r] <- 1/(1 + exp(-betaPsiRiverC[r]))
          betaPsiRiverOutD[r] <- 1/(1 + exp(-betaPsiRiverD[r]))
          for (c in 1:nCohorts){
              betaPhiRiverCohort[r,c] ~ dnorm(betaPhiRiver[r],1)
              betaPRiverCohort[r,c] ~ dnorm(betaPRiver[r],1)
              
              betaPsiRiverCohortA[r,c] ~ dnorm(betaPsiRiverA[r],1)
              betaPsiRiverCohortB[r,c] ~ dnorm(betaPsiRiverB[r],1)
              betaPsiRiverCohortC[r,c] ~ dnorm(betaPsiRiverC[r],1)
              betaPsiRiverCohortD[r,c] ~ dnorm(betaPsiRiverD[r],1)
              
              betaPhiRiverCohortOut[r,c] <- 1/(1 + exp(-betaPhiRiverCohort[r,c]))
              betaPRiverCohortOut[r,c] <- 1/(1 + exp(-betaPRiverCohort[r,c]))
              
              betaPsiRiverCohortOutA[r,c] <- 1/(1 + exp(-betaPsiRiverCohortA[r,c]))
              betaPsiRiverCohortOutB[r,c] <- 1/(1 + exp(-betaPsiRiverCohortB[r,c]))
              betaPsiRiverCohortOutC[r,c] <- 1/(1 + exp(-betaPsiRiverCohortC[r,c]))
              betaPsiRiverCohortOutD[r,c] <- 1/(1 + exp(-betaPsiRiverCohortD[r,c]))
              for (t in 1:(T-1)){
                  betaPhi[r,t,c] ~ dnorm(betaPhiRiverCohort[r,c],1)
                  betaP[r,t,c] ~ dnorm(betaPRiverCohort[r,c],1)
                  betaPhiOut[r,t,c] <- 1/(1 + exp(-betaPhi[r,t,c]))
                  betaPOut[r,t,c] <- 1/(1 + exp(-betaP[r,t,c]))
              }
          }
      }
  
      for (t in 1:(T-1)){ # loop over time
        for (c in 1:nCohorts){
          for (r in 1:(nRivers - 1)){
              lpsiA[r,t,c] ~ dnorm(betaPsiRiverCohortA[r,c], sd = 1000) # sd too big??? could group by season
              lpsiB[r,t,c] ~ dnorm(betaPsiRiverCohortB[r,c], sd = 1000)
              lpsiC[r,t,c] ~ dnorm(betaPsiRiverCohortC[r,c], sd = 1000)
              lpsiD[r,t,c] ~ dnorm(betaPsiRiverCohortD[r,c], sd = 1000)
             
              # constrain the transitions such that their sum is < 1
              psiA[r,t,c] <- exp(lpsiA[r,t,c]) / (1 + exp(lpsiA[1,t,c]) + exp(lpsiA[2,t,c]) + exp(lpsiA[3,t,c]))
              psiB[r,t,c] <- exp(lpsiB[r,t,c]) / (1 + exp(lpsiB[1,t,c]) + exp(lpsiB[2,t,c]) + exp(lpsiB[3,t,c]))
              psiC[r,t,c] <- exp(lpsiC[r,t,c]) / (1 + exp(lpsiC[1,t,c]) + exp(lpsiC[2,t,c]) + exp(lpsiC[3,t,c]))
              psiD[r,t,c] <- exp(lpsiD[r,t,c]) / (1 + exp(lpsiD[1,t,c]) + exp(lpsiD[2,t,c]) + exp(lpsiD[3,t,c]))
          }
         
          # last transition probability
          psiA[4,t,c] <- 1 - psiA[1,t,c] - psiA[2,t,c] - psiA[3,t,c]
          psiB[4,t,c] <- 1 - psiB[1,t,c] - psiB[2,t,c] - psiB[3,t,c]
          psiC[4,t,c] <- 1 - psiC[1,t,c] - psiC[2,t,c] - psiC[3,t,c]
          psiD[4,t,c] <- 1 - psiD[1,t,c] - psiD[2,t,c] - psiD[3,t,c]
       
          
          gamma[1,1,t,c] <- ilogit(betaPhi[1,t,c]) * psiA[1,t,c]
          gamma[1,2,t,c] <- ilogit(betaPhi[1,t,c]) * psiA[2,t,c]
          gamma[1,3,t,c] <- ilogit(betaPhi[1,t,c]) * psiA[3,t,c]
          gamma[1,4,t,c] <- ilogit(betaPhi[1,t,c]) * psiA[4,t,c]
          gamma[1,5,t,c] <- 1 - ilogit(betaPhi[1,t,c])
          gamma[2,1,t,c] <- ilogit(betaPhi[2,t,c]) * psiB[1,t,c]
          gamma[2,2,t,c] <- ilogit(betaPhi[2,t,c]) * psiB[2,t,c]
          gamma[2,3,t,c] <- ilogit(betaPhi[2,t,c]) * psiB[3,t,c]
          gamma[2,4,t,c] <- ilogit(betaPhi[2,t,c]) * psiB[4,t,c]
          gamma[2,5,t,c] <- 1 - ilogit(betaPhi[2,t,c])
          gamma[3,1,t,c] <- ilogit(betaPhi[3,t,c]) * psiC[1,t,c]
          gamma[3,2,t,c] <- ilogit(betaPhi[3,t,c]) * psiC[2,t,c]
          gamma[3,3,t,c] <- ilogit(betaPhi[3,t,c]) * psiC[3,t,c]
          gamma[3,4,t,c] <- ilogit(betaPhi[3,t,c]) * psiC[4,t,c]
          gamma[3,5,t,c] <- 1 - ilogit(betaPhi[3,t,c])
          gamma[4,1,t,c] <- ilogit(betaPhi[4,t,c]) * psiD[1,t,c]
          gamma[4,2,t,c] <- ilogit(betaPhi[4,t,c]) * psiD[2,t,c]
          gamma[4,3,t,c] <- ilogit(betaPhi[4,t,c]) * psiD[3,t,c]
          gamma[4,4,t,c] <- ilogit(betaPhi[4,t,c]) * psiD[4,t,c]
          gamma[4,5,t,c] <- 1 - ilogit(betaPhi[4,t,c])
          gamma[5,1,t,c] <- 0
          gamma[5,2,t,c] <- 0
          gamma[5,3,t,c] <- 0
          gamma[5,4,t,c] <- 0
          gamma[5,5,t,c] <- 1
        }
      }
       
      # gamma for the last occasion  
      for (c in 1:nCohorts){
          for (a in 1:(nRivers+1)){
              for (b in 1:nRivers){
                  gamma[a,b,T,c] <- 0
              }  
              gamma[a,5,T,c] <- 1
          }
      }
       
       
      for (i in 1:N){ # loop over individuals
          # omega for first obs      
          omega[1,1,first[i],i] <- 0          # Pr(alive A t -> non-detected t)
          omega[1,2,first[i],i] <- 1          # Pr(alive A t -> detected A t)
          omega[1,3,first[i],i] <- 0          # Pr(alive A t -> detected B t)
          omega[1,4,first[i],i] <- 0          # Pr(alive A t -> detected C t)
          omega[1,5,first[i],i] <- 0          # Pr(alive A t -> detected D t)
          omega[2,1,first[i],i] <- 0          # Pr(alive B t -> non-detected t)
          omega[2,2,first[i],i] <- 0          # Pr(alive B t -> detected A t)
          omega[2,3,first[i],i] <- 1          # Pr(alive B t -> detected B t)
          omega[2,4,first[i],i] <- 0          # Pr(alive B t -> detected C t)
          omega[2,5,first[i],i] <- 0          # Pr(alive B t -> detected C t)
          omega[3,1,first[i],i] <- 0          # Pr(alive C t -> non-detected t)
          omega[3,2,first[i],i] <- 0          # Pr(alive C t -> detected A t)
          omega[3,3,first[i],i] <- 0          # Pr(alive C t -> detected B t)
          omega[3,4,first[i],i] <- 1          # Pr(alive C t -> detected C t)
          omega[3,5,first[i],i] <- 0          # Pr(alive C t -> detected C t)
          omega[4,1,first[i],i] <- 0          # Pr(dead t -> non-detected t)
          omega[4,2,first[i],i] <- 0          # Pr(dead t -> detected A t)
          omega[4,3,first[i],i] <- 0          # Pr(dead t -> detected B t)
          omega[4,4,first[i],i] <- 0          # Pr(dead t -> detected C t)
          omega[4,5,first[i],i] <- 1          # Pr(dead t -> detected C t)
          omega[5,1,first[i],i] <- 1          # Pr(dead t -> non-detected t)
          omega[5,2,first[i],i] <- 0          # Pr(dead t -> detected A t)
          omega[5,3,first[i],i] <- 0          # Pr(dead t -> detected B t)
          omega[5,4,first[i],i] <- 0          # Pr(dead t -> detected C t)
          omega[5,5,first[i],i] <- 0          # Pr(dead t -> detected D t)
       
         
          ## DT changes:
          ## time t > first[i]:
          for(t in (first[i]+1):last[i]) {
              logit(pA[t,i]) <- betaP[1,t-1,cohort[i]]
              logit(pB[t,i]) <- betaP[2,t-1,cohort[i]]
              logit(pC[t,i]) <- betaP[3,t-1,cohort[i]]
              logit(pD[t,i]) <- betaP[4,t-1,cohort[i]]
             
              # probabilities of y(t) given z(t)
              # omega[z, y, t, i]
              
              # z=1 = alive in River 1, z=2 = alive in River 2...z=5 = dead
              # y=1 = unobserved, y=2 = observed in River 1, y=3 = observed in River 2, etc
              
              omega[1,1,t,i] <- 1 - pA[t,i]     # Pr(alive A t -> non-detected t)
              omega[1,2,t,i] <- pA[t,i]         # Pr(alive A t -> detected A t)
              omega[1,3,t,i] <- 0               # Pr(alive A t -> detected B t)
              omega[1,4,t,i] <- 0               # Pr(alive A t -> detected C t)
              omega[1,5,t,i] <- 0               # Pr(alive A t -> detected D t)
              omega[2,1,t,i] <- 1 - pB[t,i]     # Pr(alive B t -> non-detected t)
              omega[2,2,t,i] <- 0               # Pr(alive B t -> detected A t)
              omega[2,3,t,i] <- pB[t,i]         # Pr(alive B t -> detected B t)
              omega[2,4,t,i] <- 0               # Pr(alive B t -> detected C t)
              omega[2,5,t,i] <- 0               # Pr(alive B t -> detected C t)
              omega[3,1,t,i] <- 1 - pC[t,i]     # Pr(alive C t -> non-detected t)
              omega[3,2,t,i] <- 0               # Pr(alive C t -> detected A t)
              omega[3,3,t,i] <- 0               # Pr(alive C t -> detected B t)
              omega[3,4,t,i] <- pC[t,i]         # Pr(alive C t -> detected C t)
              omega[3,5,t,i] <- 0               # Pr(alive C t -> detected C t)
              omega[4,1,t,i] <- 1 - pD[t,i]     # Pr(alive D t -> non-detected t))
              omega[4,2,t,i] <- 0               # Pr(dead D t -> detected A t)
              omega[4,3,t,i] <- 0               # Pr(dead D t -> detected B t)
              omega[4,4,t,i] <- 0               # Pr(dead D t -> detected C t)
              omega[4,5,t,i] <- pD[t,i]         # Pr(alive D t -> detected D t)
              omega[5,1,t,i] <- 1               # Pr(dead t -> non-detected t)
              omega[5,2,t,i] <- 0               # Pr(dead t -> detected A t)
              omega[5,3,t,i] <- 0               # Pr(dead t -> detected B t)
              omega[5,4,t,i] <- 0               # Pr(dead t -> detected C t)
              omega[5,5,t,i] <- 0               # Pr(dead t -> detected D t)
          }
         
      } # i loop
  
      for (i in 1:N){
          y[i,first[i]:last[i]] ~ dDHMMo(init = delta[1:5],
                                         probTrans = gamma[1:5, 1:5, first[i]:last[i], cohort[i]],
                                         probObs =   omega[1:5, 1:5, first[i]:last[i], i],
                                         len = length[i],
                                         checkRowSums = 1)
      }
  
  })
  
  ##
  myConstants0 <- list(N = nrow(y),
                       T = ncol(y),
                       first = first,
                       last = last,
                       cohort = cohort,
                       nCohorts = nCohorts,
                       nRivers = nRivers,
                       season = seasonArray,
                       #                     flow = eh$flow,
                       length = last - first + 1
                       )
  
  ## DT changes:
  myData0 <- list(###yCJS = eh$eh, #y,    ## data for CJS distribution
      y = y + 1
  )   ## data for DHMM distribution
  
    ## if you change this FALSE to TRUE
  ## this makes the dataset smaller - only 200 observations,
  ## for quicker testing
  if(FALSE) {
      newN <- 20
      oldN <- dim(y)[1]
      set.seed(0)
      indToKeep <- sample(1:oldN, size = newN, replace = FALSE)
  }
  
  ## this removes fish that were only observed on the very last observation
  if(TRUE) {
      indToKeep <- which(first < ncol(y))
      newN <- length(indToKeep)
  }
  
  myConstants <- list(
      N = newN,
      T = myConstants0$T,
      first = myConstants0$first[indToKeep],
      last = myConstants0$last[indToKeep],
      nRivers = myConstants0$nRivers,
      cohort = myConstants0$cohort[indToKeep],
      nCohorts = myConstants0$nCohorts,
      #      season = myConstants$season,
      #      flow = myConstants$flow[indToKeep,],
      length = myConstants0$length[indToKeep]
  )
  
  myData <- list(
      ##yCJS = myData0$yCJS[indToKeep,],
      y = myData0$y[indToKeep,]
  )
  
    initialValues <- function(){
      list(
          betaPhiRiver = array(runif(nRivers, 0, 1), c(nRivers)),
          betaPhiRiverCohort = array(runif(nRivers * nCohorts, 0, 1), c(nRivers, nCohorts)),
          betaPhi = array(rnorm(nRivers * (myConstants$T - 1) * nCohorts , 0, 1), c(nRivers, (myConstants$T - 1), nCohorts)),
          betaPRiver = array(runif(nRivers, 0, 1), c(nRivers)),
          betaPRiverCohort = array(runif(nRivers * nCohorts, 0, 1), c(nRivers, nCohorts)),
          betaP = array(rnorm(nRivers * (myConstants$T - 1) * nCohorts , 0, 1), c(nRivers, (myConstants$T - 1), nCohorts)),
          lpsiA = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts)),
          lpsiB = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts)),
          lpsiC = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts)),
          lpsiD = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts))
          ####z = zInitsNA
      )
  }
  
  ###zInitsNA <- zInitsNA[indToKeep,]
  
  set.seed(0)
  start = Sys.time()
  ## you'll get warnings that the data 'yCJS' is not used, and the 'z' initial
  ## values are not in the model.  Those don't cause any problems,
  ## and let us use the same myData and initialValue() for both models.
  system.time(
      Rmodel <- nimbleModel(
          code = hmm.phiT_pT_psiT_DHMM,
          constants = myConstants,
          data = myData,              
          inits = initialValues(),
          calculate = FALSE
      )
  )
  
  #Rmodel$calculate()
  
  parametersToSave <- c("betaPhi", "betaPhiRiver", "betaPhiRiverCohort", 
                        "betaP",   "betaPRiver",   "betaPRiverCohort",
                        "betaPhiOut", "betaPhiRiverOut", "betaPhiRiverCohortOut", 
                        "betaPOut",   "betaPRiverOut",   "betaPRiverCohortOut",
                        
                        "psiA", "psiB", "psiC", "psiD",
                        "betaPsiRiverOutA", "betaPsiRiverOutB", "betaPsiRiverOutC", "betaPsiRiverOutD",   
                        "betaPsiRiverCohortOutA","betaPsiRiverCohortOutB","betaPsiRiverCohortOutC","betaPsiRiverCohortOutD"
                        )
  
  nIter <- 10000 #30000
  nBurnin <- 5000 #15000
  nChains <- 2
  thinRate <- 5
  
  system.time(
      conf <- configureMCMC(
          Rmodel,
          monitors = parametersToSave
      )
  )
  
  Rmcmc <- buildMCMC(conf, useConjugacy = FALSE)
  Cmodel <- compileNimble(Rmodel)
  Cmcmc <- compileNimble(Rmcmc, project = Rmodel)
  
  mcmc.phiT_pT_psiT_DHMM <- runMCMC(
      Cmcmc, 
      niter = nIter, 
      nburnin = nBurnin, 
      thin = thinRate, 
      nchains = nChains
  )
  
  end <- Sys.time()
  elapsed_phiT_pT_psiT_DHMM <- end - start
  toSave <- list(
      mcmc = mcmc.phiT_pT_psiT_DHMM, 
      elapsed = elapsed_phiT_pT_psiT_DHMM,
      name = "phiT_pT_psiT_DHMM",
      myConstants = myConstants, 
      nIter = nIter, 
      nBurnin = nBurnin,
      thinRate = thinRate, 
      nSeasons = nSeasons, 
      nCohorts = nCohorts,
      nChains = nChains
  )
  
    save(toSave, file = paste0('./models/cmrFlow4rivers/runsOut/mcmc_phiT_pT_psiT_DHMM_', substr(end,1,13), '.RData'))
    save(toSave, file = './models/cmrFlow4rivers/runsOut/mcmc_phiT_pT_psiT_DHMM_mostRecent.RData')
} else {
  load('./models/cmrFlow4rivers/runsOut/mcmc_phiT_pT_psiT_DHMM_mostRecent.RData')
}

if(plotMCMCOutput) {

  #MCMCsummary(object = mcmc.phiT_pT_cohort_flowHier, round = 2)
  #MCMCplot(object = mcmc.phiT_pT_cohort_flowHier, params = "betaPhiOut")
  MCMCplot(object = toSave$mcmc, params = "betaPhiOut")
  MCMCplot(object = toSave$mcmc, params = "betaPhiRiverOut")# 
  MCMCplot(object = toSave$mcmc, params = c("betaPhiRiverCohortOut"))
  MCMCplot(object = toSave$mcmc, params = "betaPOut")
  MCMCplot(object = toSave$mcmc, params = "betaPRiverOut")# 
  MCMCplot(object = toSave$mcmc, params = c("betaPRiverCohortOut"))
  
  MCMCplot(object = toSave$mcmc, params = "betaPsiRiverOutA")# 
  MCMCplot(object = toSave$mcmc, params = c("betaPsiRiverCohortOutA"))
  
  MCMCplot(object = toSave$mcmc, params = c("psiA"))
  
  priors <- rnorm(toSave$nIter * toSave$nChains, 0, 1/sqrt(.1))
  MCMCtrace(object = toSave$mcmc,
            #ISB = FALSE,
            #exact = TRUE, 
            params = c("betaPhiRiverOut"),
            pdf = FALSE, 
            priors = priors)
  
  priors <- runif(toSave$nIter * toSave$nChains, 0, 1)
  MCMCtrace(object = toSave$mcmc,
            #ISB = FALSE,
            #exact = TRUE, 
            params = c("betaPhiRiverCohortOut"),
            pdf = FALSE, 
            priors = priors)

}


```

### Model phiT_pT_psiT_dirch  
Phi and p and psi vary by time    
Includes transition probabilities among rivers
Added cohort (c) indexing to psi's
Fixed c,t indexing, change to t,c
Added ilogis() to betaPhi
Added hier vars on psi
Try dirchlet prior

#### Set up and run model
```{r phiT_pT_psiT_River2}

# Following https://oliviergimenez.github.io/bayesian-cr-workshop/worksheets/5_demo.html
if (rerunSurivalModels) {
  
  y <- eh$eh * eh$riverN
  (nCohorts <- nrow(unique(eh$cohorts)))
  (nSeasons <- nrow(unique(eh$seasons)))
  (nRivers <- length(unique(eh$data$riverN)))# rivers 1:4
  seasonArray <- c(3,4,1,2,3,4,1,2,3,4,1,2)
             

  first <- eh$first #apply(y, 1, function(x) min(which(x !=0)))
  last <- eh$last
  cohort = ((eh$cohorts) - min(eh$cohorts) + 1)$cohort #can't be a data frame or tibble
  
  zinits <- y + 1 # non-detection -> alive
  zinits[zinits == 2] <- 1 # dead -> alive
  zInitsNA <- ifelse(is.na(eh$flow), NA, 1)
  
  # Proportion of fish in each river on the first observation
  y1 <- y[,1]
  deltaProps <- table(y1[y1>0]) / length(y1[y1>0])
  
  #  For 3 states, we are using 4
  # -------------------------------------------------
  # Parameters for 3 sites (A, B, C):
  # phiA: survival probability site A
  # phiB: survival probability site B
  # phiC: survival probability site B
  # psiAA: movement probability from site A to site A (reference)
  # psiAB = psiA[1]: movement probability from site A to site B
  # psiAC = psiA[2]: movement probability from site A to site C 
  # psiBA = psiB[1]: movement probability from site B to site A
  # psiBB: movement probability from site B to site B (reference)
  # psiBC = psiB[2]: movement probability from site B to site C
  # psiCA = psiC[1]: movement probability from site C to site A
  # psiCB = psiC[2]: movement probability from site C to site B
  # psiCC: movement probability from site C to site C (reference)
  # pA: recapture probability site A
  # pB: recapture probability site B
  # pC: recapture probability site C
  # -------------------------------------------------
  # States (z):
  # 1 alive at A
  # 2 alive at B
  # 3 alive at C
  # 4 alive at D
  # 5 dead
  # Observations (y):  
  # 1 not seen
  # 2 seen at A 
  # 3 seen at B
  # 4 seen at C
  # 5 seen at D
  # river names: c("west brook" = 1, "wb jimmy" = 2, "wb mitchell" = 3, "wb obear" = 4)
  # -------------------------------------------------
  
  ## model code using DHMMo distribution
  hmm.phiT_pT_psiT_DHMM_dirch <- nimbleCode({
      # Initial distribution among rivers
      delta[1] <- deltaProps[1]                  # Pr(alive t = 1 and in river 1) = 0.4
      delta[2] <- deltaProps[2]
      delta[3] <- deltaProps[3]
      delta[4] <- deltaProps[4]
      delta[5] <- 0                    # Pr(dead t = 1) = 0
  
      for (r in 1:nRivers){
          betaPhiRiver[r] ~ dnorm(0,1)
          betaPRiver[r] ~ dnorm(0,1)
          
          betaPhiRiverOut[r] <- ilogit(betaPhiRiver[r])
          betaPRiverOut[r] <- ilogit(betaPRiver[r])
          
          for (c in 1:nCohorts){
              betaPhiRiverCohort[r,c] ~ dnorm(betaPhiRiver[r],1)
              betaPRiverCohort[r,c] ~ dnorm(betaPRiver[r],1)
 
              betaPhiRiverCohortOut[r,c] <- ilogit(betaPhiRiverCohort[r,c])
              betaPRiverCohortOut[r,c] <- ilogit(betaPRiverCohort[r,c])
              for (t in 1:(T-1)){
                  betaPhi[r,t,c] ~ dnorm(betaPhiRiverCohort[r,c],1)
                  betaP[r,t,c] ~ dnorm(betaPRiverCohort[r,c],1)
                  
                  betaPhiOut[r,t,c] <- ilogit(betaPhi[r,t,c])
                  betaPOut[r,t,c] <- ilogit(betaP[r,t,c])
                  
                  # move from river 'r' to one of river 1:nRivers
                  # Nice description of effect of 'alpha' on probabilities:
                  # https://stats.stackexchange.com/questions/244917/what-exactly-is-the-alpha-in-the-dirichlet-distribution
                  psi[r,1:nRivers,t,c] ~ ddirch(alpha[1:nRivers])
              }
          }
      }
  
      for (t in 1:(T-1)){ # loop over time
        for (c in 1:nCohorts){

          gamma[1,1,t,c] <- ilogit(betaPhi[1,t,c]) * psi[1,1,t,c]
          gamma[1,2,t,c] <- ilogit(betaPhi[1,t,c]) * psi[1,2,t,c]
          gamma[1,3,t,c] <- ilogit(betaPhi[1,t,c]) * psi[1,3,t,c]
          gamma[1,4,t,c] <- ilogit(betaPhi[1,t,c]) * psi[1,4,t,c]
          gamma[1,5,t,c] <- 1 - ilogit(betaPhi[1,t,c])
          gamma[2,1,t,c] <- ilogit(betaPhi[2,t,c]) * psi[2,1,t,c]
          gamma[2,2,t,c] <- ilogit(betaPhi[2,t,c]) * psi[2,2,t,c]
          gamma[2,3,t,c] <- ilogit(betaPhi[2,t,c]) * psi[2,3,t,c]
          gamma[2,4,t,c] <- ilogit(betaPhi[2,t,c]) * psi[2,4,t,c]
          gamma[2,5,t,c] <- 1 - ilogit(betaPhi[2,t,c])
          gamma[3,1,t,c] <- ilogit(betaPhi[3,t,c]) * psi[3,1,t,c]
          gamma[3,2,t,c] <- ilogit(betaPhi[3,t,c]) * psi[3,2,t,c]
          gamma[3,3,t,c] <- ilogit(betaPhi[3,t,c]) * psi[3,3,t,c]
          gamma[3,4,t,c] <- ilogit(betaPhi[3,t,c]) * psi[3,4,t,c]
          gamma[3,5,t,c] <- 1 - ilogit(betaPhi[3,t,c])
          gamma[4,1,t,c] <- ilogit(betaPhi[4,t,c]) * psi[4,1,t,c]
          gamma[4,2,t,c] <- ilogit(betaPhi[4,t,c]) * psi[4,2,t,c]
          gamma[4,3,t,c] <- ilogit(betaPhi[4,t,c]) * psi[4,3,t,c]
          gamma[4,4,t,c] <- ilogit(betaPhi[4,t,c]) * psi[4,4,t,c]
          gamma[4,5,t,c] <- 1 - ilogit(betaPhi[4,t,c])
          gamma[5,1,t,c] <- 0
          gamma[5,2,t,c] <- 0
          gamma[5,3,t,c] <- 0
          gamma[5,4,t,c] <- 0
          gamma[5,5,t,c] <- 1
        }
      }
       
      # gamma for the last occasion  
      for (c in 1:nCohorts){
          for (a in 1:(nRivers+1)){
              for (b in 1:nRivers){
                  gamma[a,b,T,c] <- 0
              }  
              gamma[a,5,T,c] <- 1
          }
      }
       
       
      for (i in 1:N){ # loop over individuals
          # omega for first obs      
          omega[1,1,first[i],i] <- 0          # Pr(alive A t -> non-detected t)
          omega[1,2,first[i],i] <- 1          # Pr(alive A t -> detected A t)
          omega[1,3,first[i],i] <- 0          # Pr(alive A t -> detected B t)
          omega[1,4,first[i],i] <- 0          # Pr(alive A t -> detected C t)
          omega[1,5,first[i],i] <- 0          # Pr(alive A t -> detected D t)
          omega[2,1,first[i],i] <- 0          # Pr(alive B t -> non-detected t)
          omega[2,2,first[i],i] <- 0          # Pr(alive B t -> detected A t)
          omega[2,3,first[i],i] <- 1          # Pr(alive B t -> detected B t)
          omega[2,4,first[i],i] <- 0          # Pr(alive B t -> detected C t)
          omega[2,5,first[i],i] <- 0          # Pr(alive B t -> detected C t)
          omega[3,1,first[i],i] <- 0          # Pr(alive C t -> non-detected t)
          omega[3,2,first[i],i] <- 0          # Pr(alive C t -> detected A t)
          omega[3,3,first[i],i] <- 0          # Pr(alive C t -> detected B t)
          omega[3,4,first[i],i] <- 1          # Pr(alive C t -> detected C t)
          omega[3,5,first[i],i] <- 0          # Pr(alive C t -> detected C t)
          omega[4,1,first[i],i] <- 0          # Pr(dead t -> non-detected t)
          omega[4,2,first[i],i] <- 0          # Pr(dead t -> detected A t)
          omega[4,3,first[i],i] <- 0          # Pr(dead t -> detected B t)
          omega[4,4,first[i],i] <- 0          # Pr(dead t -> detected C t)
          omega[4,5,first[i],i] <- 1          # Pr(dead t -> detected C t)
          omega[5,1,first[i],i] <- 1          # Pr(dead t -> non-detected t)
          omega[5,2,first[i],i] <- 0          # Pr(dead t -> detected A t)
          omega[5,3,first[i],i] <- 0          # Pr(dead t -> detected B t)
          omega[5,4,first[i],i] <- 0          # Pr(dead t -> detected C t)
          omega[5,5,first[i],i] <- 0          # Pr(dead t -> detected D t)
       
         
          ## DT changes:
          ## time t > first[i]:
          for(t in (first[i]+1):last[i]) {
              logit(pA[t,i]) <- betaP[1,t-1,cohort[i]]
              logit(pB[t,i]) <- betaP[2,t-1,cohort[i]]
              logit(pC[t,i]) <- betaP[3,t-1,cohort[i]]
              logit(pD[t,i]) <- betaP[4,t-1,cohort[i]]
             
              # probabilities of y(t) given z(t)
              # omega[z, y, t, i]
              
              # z=1 = alive in River 1, z=2 = alive in River 2...z=5 = dead
              # y=1 = unobserved, y=2 = observed in River 1, y=3 = observed in River 2, etc
              
              omega[1,1,t,i] <- 1 - pA[t,i]     # Pr(alive A t -> non-detected t)
              omega[1,2,t,i] <- pA[t,i]         # Pr(alive A t -> detected A t)
              omega[1,3,t,i] <- 0               # Pr(alive A t -> detected B t)
              omega[1,4,t,i] <- 0               # Pr(alive A t -> detected C t)
              omega[1,5,t,i] <- 0               # Pr(alive A t -> detected D t)
              omega[2,1,t,i] <- 1 - pB[t,i]     # Pr(alive B t -> non-detected t)
              omega[2,2,t,i] <- 0               # Pr(alive B t -> detected A t)
              omega[2,3,t,i] <- pB[t,i]         # Pr(alive B t -> detected B t)
              omega[2,4,t,i] <- 0               # Pr(alive B t -> detected C t)
              omega[2,5,t,i] <- 0               # Pr(alive B t -> detected C t)
              omega[3,1,t,i] <- 1 - pC[t,i]     # Pr(alive C t -> non-detected t)
              omega[3,2,t,i] <- 0               # Pr(alive C t -> detected A t)
              omega[3,3,t,i] <- 0               # Pr(alive C t -> detected B t)
              omega[3,4,t,i] <- pC[t,i]         # Pr(alive C t -> detected C t)
              omega[3,5,t,i] <- 0               # Pr(alive C t -> detected C t)
              omega[4,1,t,i] <- 1 - pD[t,i]     # Pr(alive D t -> non-detected t))
              omega[4,2,t,i] <- 0               # Pr(dead D t -> detected A t)
              omega[4,3,t,i] <- 0               # Pr(dead D t -> detected B t)
              omega[4,4,t,i] <- 0               # Pr(dead D t -> detected C t)
              omega[4,5,t,i] <- pD[t,i]         # Pr(alive D t -> detected D t)
              omega[5,1,t,i] <- 1               # Pr(dead t -> non-detected t)
              omega[5,2,t,i] <- 0               # Pr(dead t -> detected A t)
              omega[5,3,t,i] <- 0               # Pr(dead t -> detected B t)
              omega[5,4,t,i] <- 0               # Pr(dead t -> detected C t)
              omega[5,5,t,i] <- 0               # Pr(dead t -> detected D t)
          }
         
      } # i loop
  
      for (i in 1:N){
          y[i,first[i]:last[i]] ~ dDHMMo(init = delta[1:5],
                                         probTrans = gamma[1:5, 1:5, first[i]:last[i], cohort[i]],
                                         probObs =   omega[1:5, 1:5, first[i]:last[i], i],
                                         len = length[i],
                                         checkRowSums = 1)
      }
  
  })
  
  ##
  myConstants0 <- list(N = nrow(y),
                       T = ncol(y),
                       first = first,
                       last = last,
                       cohort = cohort,
                       nCohorts = nCohorts,
                       nRivers = nRivers,
                       season = seasonArray,
                       #                     flow = eh$flow,
                       length = last - first + 1,
                       alpha = c(1,1,1,1)
                       )
  
  ## DT changes:
  myData0 <- list(###yCJS = eh$eh, #y,    ## data for CJS distribution
      y = y + 1
  )   ## data for DHMM distribution
  
    ## if you change this FALSE to TRUE
  ## this makes the dataset smaller - only 200 observations,
  ## for quicker testing
  if(FALSE) {
      newN <- 20
      oldN <- dim(y)[1]
      set.seed(0)
      indToKeep <- sample(1:oldN, size = newN, replace = FALSE)
  }
  
  ## this removes fish that were only observed on the very last observation
  if(TRUE) {
      indToKeep <- which(first < ncol(y))
      newN <- length(indToKeep)
  }
  
  myConstants <- list(
      N = newN,
      T = myConstants0$T,
      first = myConstants0$first[indToKeep],
      last = myConstants0$last[indToKeep],
      nRivers = myConstants0$nRivers,
      cohort = myConstants0$cohort[indToKeep],
      nCohorts = myConstants0$nCohorts,
      #      season = myConstants$season,
      #      flow = myConstants$flow[indToKeep,],
      length = myConstants0$length[indToKeep],
      alpha = myConstants0$alpha
  )
  
  myData <- list(
      ##yCJS = myData0$yCJS[indToKeep,],
      y = myData0$y[indToKeep,]
  )
  
    initialValues <- function(){
      list(
          betaPhiRiver = array(runif(nRivers, 0, 1), c(nRivers)),
          betaPhiRiverCohort = array(runif(nRivers * nCohorts, 0, 1), c(nRivers, nCohorts)),
          betaPhi = array(rnorm(nRivers * (myConstants$T - 1) * nCohorts , 0, 1), c(nRivers, (myConstants$T - 1), nCohorts)),
          betaPRiver = array(runif(nRivers, 0, 1), c(nRivers)),
          betaPRiverCohort = array(runif(nRivers * nCohorts, 0, 1), c(nRivers, nCohorts)),
          betaP = array(rnorm(nRivers * (myConstants$T - 1) * nCohorts , 0, 1), c(nRivers, (myConstants$T - 1), nCohorts)),
          psi = array(1/nRivers, c(nRivers, nRivers, (myConstants$T - 1), nCohorts))
          # lpsiA = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts)),
          # lpsiB = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts)),
          # lpsiC = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts)),
          # lpsiD = array(rnorm((nRivers - 1) * (myConstants$T - 1) * nCohorts, 0, 1), c((nRivers - 1), (myConstants$T - 1), nCohorts))
          ####z = zInitsNA
      )
  }
  
  ###zInitsNA <- zInitsNA[indToKeep,]
  
  set.seed(0)
  start = Sys.time()
  ## you'll get warnings that the data 'yCJS' is not used, and the 'z' initial
  ## values are not in the model.  Those don't cause any problems,
  ## and let us use the same myData and initialValue() for both models.
  system.time(
      Rmodel <- nimbleModel(
          code = hmm.phiT_pT_psiT_DHMM_dirch,
          constants = myConstants,
          data = myData,              
          inits = initialValues(),
          calculate = FALSE
      )
  )
  
  #Rmodel$calculate()
  
  parametersToSave <- c("betaPhi", "betaPhiRiver", "betaPhiRiverCohort", 
                        "betaP",   "betaPRiver",   "betaPRiverCohort",
                        "betaPhiOut", "betaPhiRiverOut", "betaPhiRiverCohortOut", 
                        "betaPOut",   "betaPRiverOut",   "betaPRiverCohortOut",
                        
                        "psi"
                        )
  
  nIter <- 5000 #30000
  nBurnin <- 2500 #15000
  nChains <- 2
  thinRate <- 5
  
  system.time(
      conf <- configureMCMC(
          Rmodel,
          monitors = parametersToSave
      )
  )
  
  Rmcmc <- buildMCMC(conf, useConjugacy = FALSE)
  Cmodel <- compileNimble(Rmodel)
  Cmcmc <- compileNimble(Rmcmc, project = Rmodel)
  
  mcmc.phiT_pT_psiT_DHMM_dirch <- runMCMC(
      Cmcmc, 
      niter = nIter, 
      nburnin = nBurnin, 
      thin = thinRate, 
      nchains = nChains
  )
  
  end <- Sys.time()
  elapsed_phiT_pT_psiT_DHMM_dirch <- end - start
  toSave <- list(
      mcmc = mcmc.phiT_pT_psiT_DHMM_dirch, 
      elapsed = elapsed_phiT_pT_psiT_DHMM_dirch,
      name = "phiT_pT_psiT_DHMM_dirch",
      myConstants = myConstants, 
      nIter = nIter, 
      nBurnin = nBurnin,
      thinRate = thinRate, 
      nSeasons = nSeasons, 
      nCohorts = nCohorts,
      nChains = nChains
  )
  
    save(toSave, file = paste0('./models/cmrFlow4rivers/runsOut/mcmc_phiT_pT_psiT_DHMM_dirch_', substr(end,1,13), '.RData'))
    save(toSave, file = './models/cmrFlow4rivers/runsOut/mcmc_phiT_pT_psiT_DHMM_dirch_mostRecent.RData')
} else {
  load('./models/cmrFlow4rivers/runsOut/mcmc_phiT_pT_psiT_DHMM_dirch_mostRecent.RData')
}

if(plotMCMCOutput) {

  #MCMCsummary(object = mcmc.phiT_pT_cohort_flowHier, round = 2)
  #MCMCplot(object = mcmc.phiT_pT_cohort_flowHier, params = "betaPhiOut")
  MCMCplot(object = toSave$mcmc, params = "betaPhiOut")
  MCMCplot(object = toSave$mcmc, params = "betaPhiRiverOut")# 
  MCMCplot(object = toSave$mcmc, params = c("betaPhiRiverCohortOut"))
  MCMCplot(object = toSave$mcmc, params = "betaPOut")
  MCMCplot(object = toSave$mcmc, params = "betaPRiverOut")# 
  MCMCplot(object = toSave$mcmc, params = c("betaPRiverCohortOut"))
  
  MCMCplot(object = toSave$mcmc, params = "betaPsiRiverOutA")# 
  MCMCplot(object = toSave$mcmc, params = c("betaPsiRiverCohortOutA"))
  
  MCMCplot(object = toSave$mcmc, params = c("psi"))
  
  priors <- rnorm(toSave$nIter * toSave$nChains, 0, 1/sqrt(.1))
  MCMCtrace(object = toSave$mcmc,
            #ISB = FALSE,
            #exact = TRUE, 
            params = c("betaPhiRiverOut"),
            pdf = FALSE, 
            priors = priors)
  
  priors <- runif(toSave$nIter * toSave$nChains, 0, 1)
  MCMCtrace(object = toSave$mcmc,
            #ISB = FALSE,
            #exact = TRUE, 
            params = c("betaPhiRiverCohortOut"),
            pdf = FALSE, 
            priors = priors)

  priors <- runif(toSave$nIter * toSave$nChains, 0, 1)
  MCMCtrace(object = toSave$mcmc,
            #ISB = FALSE,
            #exact = TRUE, 
            params = c("psi"),
            pdf = FALSE, 
            priors = priors)
  
}


```








