# Young-of-year Growth Model {#yoyGrowthModel}
```{r globalYOY, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, out.width = '100%')
```

```{r librariesYOY, echo = FALSE}
library(tidyverse)
load(file = './data/cdWB_electro.RData')
```

The question here is what is driving _body size variation_ across years in brook trout and brown trout in the WB?  

We focus on ageInSamples == 1 (age-0 fish in the fall sample) fish for growth model. This is the first sampling occasion that most fish are big enough to tag. Not all fish are big enough, however, and there is a number of untagged fish each year. We need to include both tagged and untagged fish in our age-0 size model.

Factors to include in the model are 
1. Sample date  
2. Cumulative temperature prior to sampling  
3. Cumulative flow prior to sampling  
4. Extreme flow events?  
5. Fish density, maybe total and age-0 across all three salmonids  

## Raw data
test

### Brook Trout, West brook  
```{r raw data plots1}
  cd1 <- cdWB_electro %>% filter(ageInSamples == 1, species != 'ats')

  plotSppRiv = function(s, r) { 
    ggplot(cd1 %>% filter(species == s, riverOrdered == r), aes(observedLength, color = is.na(tag))) +
      geom_freqpoly() +
      geom_vline(xintercept = 60) +
      ggtitle(paste(s, r, sep = ', ')) +
      xlim(c(30,125)) +
      facet_wrap(~ year, scales = "free_y")
  }
  
  species = 'bkt'
  riverOrdered = "west brook"

  plotSppRiv(species, riverOrdered)
```  

### Brook Trout, wb jimmy 
```{r raw data plots2, echo = FALSE}
  species = 'bkt'
  riverOrdered = "wb jimmy"

  plotSppRiv(species, riverOrdered)
```  

### Brook Trout, wb mitchell  
```{r raw data plots3, echo = FALSE}
  species = 'bkt'
  riverOrdered = "wb mitchell"

  plotSppRiv(species, riverOrdered)
```  

### Brook Trout, wb obear  
```{r raw data plots4, echo = FALSE}
  species = 'bkt'
  riverOrdered = "wb obear"

  plotSppRiv(species, riverOrdered)
```  

### Brown Trout, West brook  
```{r raw data plots5, echo = FALSE}
  species = 'bnt'
  riverOrdered = "west brook"

  plotSppRiv(species, riverOrdered)
```  

### Brown Trout, wb jimmy 
```{r raw data plots6, echo = FALSE}
  species = 'bnt'
  riverOrdered = "wb jimmy"

  plotSppRiv(species, riverOrdered)
```  

### Brown Trout, wb mitchell  
```{r raw data plots7, echo = FALSE}
  species = 'bnt'
  riverOrdered = "wb mitchell"

  plotSppRiv(species, riverOrdered)
```  

### Brown Trout, wb obear - there are no Brown trout in O'Bear  

### Trout, in the WB mainstem only
```{r}
  cdWB1 <- cdWB_electro %>% filter(ageInSamples == 1, species != 'ats', riverOrdered == "west brook")

  ggplot(cdWB1, aes(observedLength)) +
    geom_freqpoly() +
    geom_vline(xintercept = 60, color = 'orange') +
    facet_grid(species ~ year)
  
  ggplot(cdWB1, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60) +
    facet_grid(species ~ year)
```
  

### Why are there untagged fish bigger than 60mm?  
Check 2002/bkt/WB, as an example

```{r 2002 untagged}

  cdWB12002BKT <- cdWB1 %>% filter(year == 2002, species == "bkt")
  table(is.na(cdWB12002BKT$tag))
  
  ggplot(cdWB12002BKT, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60)
  
  cdWB12002BKT %>% filter(observedLength > 60, is.na(tag)) %>% data.frame()
  
  # looks like untagged area=inside fish wee morts, the rest were above or below
  ggplot(cdWB12002BKT, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60) +
    facet_grid(~area)
  
  # check 2003
  ggplot(cdWB1 %>% filter(year == 2003, species == "bkt"), aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60) +
    facet_grid(~area)
  # same story as 2002
```

### Why no untagged fish at all for 2000 and 2001?

```{r untagged 2000, 2001}

  cdWB12000_2001BKT <- cdWB1 %>% filter(year %in% 2000:2001, species == "bkt")
  table(is.na(cdWB12000_2001BKT$tag))
  table(cdWB12000_2001BKT$observedLength)

  cdWB12000_2001BKT %>% filter(observedLength < 60)
  
  ggplot(cdWB12000_2001BKT, aes(observedLength, color = is.na(tag))) +
    geom_freqpoly() +
    geom_vline(xintercept = 60)
```

## Yearly means
```{r means}

means <- cdWB1 %>% group_by(species, year) %>% summarize(meanLength =
mean(observedLength, na.rm = TRUE), meanYday = mean(yday, na.rm = TRUE))

ggplot(means, aes(year, meanLength, color = species)) + geom_point() +
geom_line()

ggplot(means, aes(meanYday, meanLength, color = species)) +
geom_point() + geom_smooth(method = "lm")
```





